### 2024-11-26 paper/paper.qmd
intro, abstract, title completed

---
title: "Does the Price Stability of 'No Name' Mean No Price Increases? A Strategy by Loblaws to Retain Customers"
subtitle: "Price Changes of Private-Label Brands: Stability, Strategy, and Consumer Impact"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/ChenmingZ2000/Loblaws-NoName-Grocery-Price-Trend-Analysis.git](https://github.com/ChenmingZ2000/Loblaws-NoName-Grocery-Price-Trend-Analysis.git).
        We gratefully acknowledge Jacob Filipp’s groceries dataset used in this analysis. The data is available at: [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/)"
date: today
date-format: long
abstract: "This paper examines the pricing dynamics of Loblaws' private-label brand, No Name, in the context of rising grocery prices and consumer dissatisfaction in Canada. Using a hierarchical Bayesian model, we analyzed price fluctuations of No Name products across multiple categories to identify trends and potential strategies employed by Loblaws. Our findings shows that while prices exhibit minor fluctuations, No Name products have largely remained stable, suggesting a deliberate strategy by Loblaws to retain customers amidst growing competition and economic pressures. This study highlights the role of private-label brands in maintaining consumer loyalty and offers ideas about how large retailers balance affordability and profitability in an inflationary environment."
format: pdf
number-sections: true
bibliography: references.bib
pdf-engine: lualatex
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(bayesplot)
library(patchwork)
library(arrow)
```

# Introduction

The Bank of Canada has noted that, since the economy began recovering from the pandemic, Canadians have experienced high inflation across many goods and services, with food prices rising faster than most other categories. Several factors have contributed to the surge in food costs, including supply chain disruptions, labor shortages, adverse weather conditions, and geopolitical events [see more details in @bankofcanada2024groceries and @statcan2022foodprices]. For example, in September 2022, annual food price inflation at grocery stores reached 11.4%, the fastest pace since 1981. In Toronto, food bank visits also set a new record, with 3.49 million visits recorded between April 2023 and April 2024, marking a significant increase from the previous year. This trend highlights the growing difficulty residents face in coping with the rising cost of living [see more details in @cbc2024foodbank].

Among Canada’s major grocery vendors, Loblaw Companies Limited has been at the center of widespread discussions and dissatisfaction due to price increases. As reported in @cbc2024people consumers have expressed frustration with the continued rise in food prices, particularly as inflation has begun to slow, raising questions about the pricing power of large grocery chains. Furthermore, Loblaw and its parent company, George Weston Limited, agreed to pay CAD 500 million to settle a class-action lawsuit related to bread price-fixing allegations, further exacerbating public dissatisfaction with the company's pricing strategies [see more details in @cbc2024].

The estimand of this study is the price fluctuations of grocery items under Loblaw’s private-label brand, No Name. As one of Canada’s largest grocery retailers, Loblaw’s No Name brand has a high level of market coverage and consumer recognition nationwide. Positioned as an affordable, high-value brand, No Name aligns closely with the needs of consumers facing inflation and rising food prices. It also serves as a key tool for Loblaw to increase customer retention. By studying No Name's pricing and market behavior, we aim to infer Loblaw’s overall pricing strategies.
Our analysis indicates that while most No Name products exhibit price fluctuations, their prices have remained relatively stable without significant increases. This suggests that Loblaw is likely maintaining control over No Name prices and using the PC Optimum loyalty program to prevent customer attrition.

The remainder of this paper is structured as follows. @sec-data introduces the data sources, preprocessing steps, variable selection criteria, and potential relationships between independent and dependent variables. @sec-model provides a concise explanation of the hierarchical Bayesian model, including the design rationale and the selection of priors and random effects, with more details on the model’s posterior in [Appendix -@sec-model-details]. @sec-result presents the main findings of the study, including a summary of the model output to support its accuracy and predictions of grocery price trends by different categories of “No Name” products through February 15, 2025. Finally, @sec-discussion includes the implications of the results, limitations of the data, complexities of incorporating real-world factors into the dataset, potential future pricing strategies by Loblaw, and steps for further research.

# Data {#sec-data}

## Overview

The product pricing data for the "No Name" brand sold by Loblaws, used in this analysis, is included in the dataset provided by [@hammer]. This dataset compiles detailed product information, including vendors, product names, brands, categories, and unit prices. Additionally, it includes time-series pricing records for each product, capturing timestamps of price changes, current prices, unit prices, and quantities sold. These records are periodically updated on the Hammer platform to reflect the latest pricing trends. This dataset forms the foundation for analyzing price trends, identifying unit variations, and categorizing products into specific types for further analysis. @tbl-view previews the information for the target products included in this analysis, starting from June 2024.

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr], `arrow` [@arrow], `brms` [@brms_package] `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

## Measurement
In the real world, pricing data for consumer goods is a dynamic reflection of market trends, regional preferences, and competitive strategies. The prices of products fluctuate based on supply chain disruptions, seasonal demand, and retail strategies, making it essential to track and analyze this information systematically. The Hammer dataset serves as a resource in understanding these trends, offering a detailed collection of product pricing data sourced from eight vendors. This dataset is embedded with complexities arising from the data collection process, competitive pricing strategies, and evolving consumer behavior.

Based on content on the Project Hammer website, the data collection process for @hammer is web scraping. This is one of the typical approaches used to construct structured datasets from online retail platforms. That is, retailers provide public access to their product inventories and prices through their websites, making them prime candidates for automated data extraction. Web scraping tools are deployed to extract product details such as names, brands, vendors, categories, unit prices, and quantities, while timestamps capture the exact moment of data retrieval to ensure time-sensitive accuracy. However, even with automated methods, the process is not without its challenges. Retailers often introduce barriers like CAPTCHA systems or dynamic content loading, which require solutions such as headless browsers or API integrations to ensure smooth data acquisition.

Sampling and representativeness are considerations in Hammer's data collection process.Inherent biases may exist. For instance, not all vendors update their online product catalogs in real time, which may introduce lag in capturing price changes or include the same data for different products. Once the raw data is collected, it undergoes cleaning and standardization. Product names, for instance, often include variations due to marketing strategies (e.g., "Club Size" or "Family Pack"), which are standardized to ensure consistency across entries. Similarly, data with missing values—such as prices or unit information—are flagged for further investigation.

Finally, the cleaned and standardized data is stored in a structured database, organized by product ID, timestamps, and categorical variables such as vendor, brand, and product category. This structured approach allows researchers to track pricing trends over time, analyze variations across regions or vendors, and categorize products into specific types. Hammer's database not only facilitates trend analysis but also enables broader comparisons and aggregations, allowing researchers to examine shifts in consumer behavior or retailer strategies.

## Analysis Data

This analysis focuses on Loblaws' "No Name" brand products, with the goal of understanding pricing trends. We only selected observations where the vendor is Loblaws and the brand is "No Name" because Loblaws announced a price freeze on more than 1,500 "No Name" products until January 31, 2023, to address the pressures of rising food prices. However, with the price freeze period now over in 2024, we are particularly interested in analyzing the price changes of "No Name" products [see more details in @loblaw2022]. At the same time, we prioritized high-quality data by filtering out records with missing or incomplete price information, as we have plenty of price records for a product in one day and unreliable data could introduce significant bias and reduce the accuracy of our trend analysis. In researching price trends and product variations, we selected specific variables that are impactful for analyzing pricing behavior affected by product categorization. Below is an overview of the data used for this analysis @tbl-view, along with the selected outcome and predictor variables.

### Outcome Variables

In this analysis of Loblaws' "No Name" brand products 2024, the primary outcome variable is the `current_price`, which represents the retail price of a product at the time of the record. Understanding the pricing trends of these products allows us to analyze consumer affordability, retailer pricing strategies, and competitive positioning in the market. By focusing on `current_price`, we aim to identify influences that influence price fluctuations over time and assess how product size, unit and type variations contribute to these trends. This metric serves as the foundation for evaluating the dynamics of product pricing and finding differences across product categories.

### Predictor Variables
In this analysis, the predictor variables are the variables that we believe can influence the current price of products in the Loblaws grocery dataset. These variables are chosen based on their potential to explain variations in product pricing. Below are the predictor variables included in our model, along with an explanation of their significance.

**Size**: Product size is a determinant of pricing, as larger sizes generally cost more due to the increased quantity. However, pricing strategies such as bulk discounts or price-per-unit variations can create non-linear relationships between size and price, making size an important factor to analyze.

**Type**: Product type (e.g., snacks, seasoning, dairy) reflects category-level differences in pricing strategies. Different types of products may have varying production costs, supply chain logistics, and market demand, all of which contribute to price fluctuations.

**Date**: Time plays an important role in grocery pricing due to reasons like seasonal demand, inflation, supply chain disruptions, and promotions. Analyzing the impact of time allows us to capture temporal trends and shifts in pricing behavior, such as rising costs during inflationary periods or discounts during specific seasons.

### Exclusions and Considerations
Other variables (e.g., vendor or brand-specific promotions) could also influence pricing. However, in this analysis, we focus specifically on the effects of `size`, `type`, and `date`, as they represent the most consistent and accessible predictors in the dataset. Additional variables, while potentially relevant, may not be consistently available or may introduce unnecessary complexity into the model. For example, promotions or regional pricing differences were not included in this analysis due to large amounts of data unavailability. Additionally, while `old_price` might influence pricing strategies, this dataset does not provide enough details. By focusing on the variables above, we aim to construct a reliable and interpretable model for understanding price dynamics within Loblaws' "No Name" brand products.

```{r, message=FALSE}
#| echo: false
#| eval: true
#| warning: false
#| message: false

loblaws_data <- read_parquet(here::here("data/02-analysis_data/loblaws_data.parquet"))
```

```{r}
#| label: tbl-view
#| tbl-cap: The First Few Lines of Cleaned Grocery Price Analysis Data
#| echo: false

sorted_data <- loblaws_data %>%
  arrange(date)

sorted_data <- sorted_data %>%
  select(product_id, date, current_price, product_name, type, size)

kable(head(sorted_data, 7), 
      col.names = c("product_id", "date", "current_price", "product_name", "type", "size"), 
      align = c("l", "c", "c", "c", "c", "c")) %>%
kable_styling(font_size = 7)
```

### Relations Between Variables

The following figures, @fig-1 and @fig-2, present various aspects of data related to Loblaws' "No Name" grocery prices, helping us understand how item prices change with different variables. In the left panel of @fig-1 and both panels in @fig-2, the x-axis is marked with dates categorized by month. The chart analyzes the relationships between product prices, size, type, and date, showing how these variables may interact and influence pricing trends. Product size and type have an impact on pricing trends, and we anticipate that incorporating these variables into the predictive model will improve the accuracy of price predictions under different product and market conditions.

While we observe certain significant fluctuations in the right panel of @fig-2, which may indicate abnormal prices being recorded or the presence of products with extreme prices in specific categories during certain periods, we cannot directly remove this data as it still contains information we need. In other words, the prices of these products can also explain some of the reasons behind price fluctuations. Completely removing these observations would weaken the model's ability to reflect price trends accurately. Therefore, we choose to retain these observations and mitigate the biases they introduce by combining other variables to build a more robust predictive model.

The boxplot on the right part of @fig-1, shows the pricing differences across product sizes. Compared to "Small" and "Large" products, "Extra Large" products have a higher median price and a wider interquartile range. This reflects inherent pricing strategies where larger quantities are priced higher, but not always linearly (e.g., discounts for bulk purchases). However, the presence of outliers, especially in the "Extra Large" and "Large" categories, indicates that even within the same size group, different product types or brands can significantly impact final pricing. The narrow boxplot for "Small" products suggests a smaller range of price variation, which aligns with the lack of price fluctuations observed in the left panel of @fig-1 for small-sized products.

@fig-2 shows how average prices vary over time, segmented by product `type` (left) and `size` (right). In the left panel, we observe distinct pricing patterns across product types. For instance, "Seasoning" demonstrates higher volatility compared to "Flour/Rice," likely due to differences in demand elasticity. The "Milk" category shows occasional price dips, possibly influenced by the introduction of new products or promotional events.The right panel shows price trends segmented by size. "Extra Large" products generally have higher average prices but exhibit more unstable trends over time. In contrast, "Small" products show minimal price variation, possibly due to their inherently lower pricing or because their unit price is relatively higher compared to larger sizes.

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=3}
#| label: fig-1
#| fig-cap: Price Trend for Six of the Loblaws "No Name" Groceries Over Time (left) and Price Distribution by Size (right).
#| echo: false

# Plot price trends over time by product_id
selected_ids <- c("1870038", "1876096", "1871388", "1295620", "1876747", 
                  "1870410")

filtered_data <- loblaws_data %>%
  filter(product_id %in% selected_ids) %>%
  mutate(label = paste(product_name, size, sep = ", "))

p1 <- ggplot(filtered_data, aes(x = date, y = current_price, group = product_id, color = label)) +
  geom_line(alpha = 0.7, size = 1) +
  labs(x = "Date", 
       y = "Price of the Day (CAD)", 
       title = "Price Trend for Products by Date",
       subtitle = "Product randomly selected",
       color = "Product Name and Size") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

p2 <- ggplot(loblaws_data, aes(x = size, y = current_price, fill = size)) +
  geom_boxplot() +
  labs(x = "Size", y = "Price (CAD)", title = "Price Distribution by Size")  +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

p1 + p2
```

```{r, message=FALSE, warning=FALSE, fig.width= 7, fig.height= 3}
#| label: fig-2
#| fig-cap: Average Price Trend for Loblaws "No Name" Groceries Over Time Categorized in Type (left) and in Size (right).
#| echo: false
# Group data by date and product type, and calculate average price for each type per date
type_trend_data <- loblaws_data %>%
  group_by(date, type) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

p3 <- ggplot(type_trend_data, aes(x = date, y = avg_price, color = type)) +
  geom_line(size = 0.7) +
  labs(
    x = "Date",
    y = "Average Price (CAD)",
    title = "Price Trend by Time and Product Type",
    color = "Product Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

# Group data by date and product size, and calculate average price for each size per date
size_trend_data <- loblaws_data %>%
  group_by(date, size) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

# Plot the trend
p4 <- ggplot(size_trend_data, aes(x = date, y = avg_price, color = size)) +
  geom_line(size = 0.7) +
  labs(
    x = "Date",
    y = "Average Price (CAD)",
    title = "Price Trend by Time and Product Size",
    color = "Product Size"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

p3 + p4
```

# Model {#sec-model}
Our model references the analytical approach proposed by @gamma_2018, adopting a hierarchical Bayesian framework to model the price trends of Loblaws' "No Name" grocery products. Specifically, the model is based on a Gamma regression framework with a log-link function, designed to handle the positively skewed distribution of product prices. By using the Gamma distribution, we can capture the inherent variability in prices while adhering to the non-negative nature of price data. The Gamma distribution is particularly well-suited for modeling continuous, non-negative, and positively right-skewed data, which are common in economic contexts such as insurance claims and survival data [@gamma_2018].

The goal of our modeling strategy is to build a pricing model aimed at predicting changes in `current_price` over time while considering key variables that influence price fluctuations. The structure of the model combines fixed effects and random effects and includes multiple variables, such as `product_id`, product `size`, and product `type`. This modeling approach is designed to balance the complexity of the data while ensuring robust predictions. The hierarchical structure of the model allows us to capture variations at the level of individual products while identifying overall trends across different product sizes and types. It also enables us to forecast future price trends of Loblaws "No Name" groceries, observing whether the retailer's pricing behavior aligns with their commitments. Here, we provide a brief overview of the Bayesian analytical model used to investigate the price trends of Loblaws "No Name" products. Background details, model assumptions, and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

This model is a generalized linear mixed model (GLMM) within a Bayesian framework, specifically designed to handle positively skewed data:

$y_i \sim \text{Gamma}(\mu_i, \phi)$

Define $y_i$ as the current price of the product in the i-th observation, assumed to follow a Gamma distribution with mean $\mu_i$ and shape parameter $\phi$. This model accounts for variations across different product sizes, types, and temporal trends, with random intercepts for each product ID.

The linear predictor for the mean ($\mu_i$) of the Gamma distribution is defined as follows:

$log(\mu_i)$ = $\alpha$ + $\beta_1 \cdot \text{size}_ i$ + $\beta_2 \cdot \text{type}_i$ + $\beta_3 \cdot \text{date}_i$ + $u_{\text{productID}[i]}$

where:

- $y_i$ is the observed current price for the i-th product.

- $\mu_i$ is the predicted mean price for the i-th product, modeled on a log scale.

- $\alpha \sim \text{Normal}(2, 5)$ is the global intercept representing the baseline price.

- $\beta_1 \sim \text{Normal}(0, 5)$ captures the effect of product size on pricing (e.g., Small, Large, Extra Large).

- $\beta_2 \sim \text{Normal}(0, 5)$ represents the effect of product type (e.g., Snacks, Flour/Rice, Cooking Oil) on pricing.

- $\beta_3 \sim \text{Normal}(0, 5)$ accounts for the temporal trend in prices over time.

Random Effects: $u_{\text{productID}} \sim \text{Normal}(0, \sigma_{\text{productID}})$

where $u_{\text{productID}[i]}$ is the random intercept for each product ID, capturing product-specific variations in pricing.

## Model justification
### Model Prior Selection, Assumptions, and Validation
Our prior selection was informed by the research in gamma modeling @Veazie2023GammaModel, we aim to provide sufficient flexibility for the model to adapt to the data. For example, the prior for the intercept was set as $\text{Normal}(2, 5)$, reflecting the assumption that the baseline log-transformed price of Loblaws "No Name" products fluctuates around 2 while allowing for some uncertainty. For the effect coefficients of variables such as `size`, `type`, and `date`, we chose priors of $\text{Normal}(0, 5)$, representing the expectation of moderate impacts of these variables on pricing. These priors enable the model to adjust based on observed data.

The model assumes that price variations at the product level are non-negative, right-skewed, and that the price differences associated with size and type are random and normally distributed. While these assumptions simplify computation and enhance interpretability, they may not fully capture the impact of market conditions, regional competition, or seasonal trends on pricing dynamics. For instance, the model assumes that the effect of size is consistent across all product types, but in reality, certain product types may be more significantly influenced by supply chain issues or promotional activities. Additionally, the model incorporates random effects for individual products to capture specific variations for each product, assuming these effects remain consistent over time.

The model was implemented using the `rstanarm` package and employed MCMC sampling for Bayesian inference. To validate the model, we conducted out-of-sample testing and calculated the predictive accuracy using RMSE (Root Mean Square Error), which was 0.83 on the log-transformed price scale. In the context of grocery price data, this level of error is acceptable, as market forces and external factors inherently introduce noise. Convergence diagnostics (e.g., $\hat{R}$) showed values close to 1 for all parameters, indicating successful convergence of the chains. The effective sample size (n_eff) for key parameters exceeded 500, ensuring the robustness of the inferences. Sensitivity analysis on priors revealed that minor adjustments to prior specifications did not significantly affect the model's predictions, demonstrating high model stability.

We also compared this model to alternative approaches. A simpler linear regression model without random effects failed to account for product-level variations, particularly for products with fewer observations, resulting in significantly higher prediction errors. Conversely, a more complex model incorporating additional factors such as promotional pricing or competitor prices achieved only marginal improvements in predictive accuracy but at a substantial increase in computational cost, making it overly complex. Therefore, the current model, which integrates a hierarchical structure and focuses on size, type, and temporal trends, strikes a good balance between predictive accuracy, interpretability, and computational efficiency.

### Expectation of relationships between variables

For the variable `type`, we anticipate that different product types will influence the estimation of current prices. While we cannot precisely determine which product categories incur higher costs due to supply chain or consumer demand, it is reasonable to assume that categories such as "Snacks" or "Flour/Rice" may exhibit distinct pricing dynamics. For instance, "Snacks" might experience greater price volatility due to internet trends or elastic demand, whereas "Flour/Rice" could reflect a more stable trend. These category-specific variables introduce variability into the overall pricing estimates.

For the variable `date`, we hypothesize that the prices of Loblaws "No Name" products fluctuate over time, reflecting seasonal trends, market dynamics, and promotional activities. This hypothesis is based on historical observations, where prices often vary due to promotions or stock-outs. Long-term trends, on the other hand, may reflect broader economic conditions or changes in production costs.

The variable size captures the impact of product quantity on pricing. We expect larger-sized products, such as "Extra Large," to generally have higher prices due to increased quantities. However, economies of scale may result in lower unit prices for larger sizes, while smaller sizes, despite lower overall costs, may have higher unit prices. This pricing strategy is common in retail to ensure accessibility for different consumer groups while optimizing profit margins.

To account for variations among individual products, we introduced a random intercept `(1 | product_id)` for each product. This intercept captures product-specific price differences, such as brand effects, unique demand characteristics, or production costs.

By modeling the interaction between size and type `(size * type)`, the model allows for a more detailed understanding of relationships, such as pricing differences for "Extra Large" products between "Cooking Oil" and "Milk." This interaction term helps identify combined effects that cannot be fully explained by individual variables. Together, these variables enable the model to capture the dynamic pricing structure of Loblaws "No Name" products, reflecting time trends and structural differences across product categories and sizes.

# Results {#sec-result}

## Model Output

Our model results are summarized in @tbl-model. This summary shows the effects of product attributes, including size and type, on price predictions. For instance, products categorized as "Large" are associated with a price reduction of 0.47 CAD compared to the baseline (Extra Large), while "Small" products show an even larger reduction of 0.98 CAD. Similarly, "Snacks" exhibit a negative effect of 0.39 CAD on price compared to the baseline type (Cooking Oil), while "Flour/Rice" shows a negligible effect. 
The model achieves an $R^2$ value of 0.967, indicating that it explains 96.7% of the variation in the data. This shows that the model effectively captures the main patterns in the data. The marginal $R^2$, which excludes the random effects, is 0.742, suggesting that fixed effects alone explain a substantial portion of the variance. The Intraclass Correlation Coefficient (ICC) is 0.1, indicating that only 10% of the variability is attributable to clustering by `product_id`, suggesting relatively low within-group correlation. Additional metrics support the model's robustness. The WAIC (6585.0) and LOOIC (6585.1) values confirm that the model balances predictive accuracy with generalizability, avoiding overfitting.

```{r, message=FALSE}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <- readRDS(file = here::here("models/price_model.rds"))

prediction_data <- read_parquet(here::here("data/02-analysis_data/predicted_prices.parquet"))

```

```{r, message=FALSE}
#| echo: false
#| eval: true
#| label: tbl-model
#| tbl-cap: Model summary table for model which predicts Loblaws "No Name" grocery price in 2024
#| warning: false
#| message: false

modelsummary::modelsummary(
  list("Loblaws 'No Name' Price Gamma model" = first_model),
  statistic = "mad",
  fmt = 2,
  coef_map = c("Sigma[product_id × (Intercept),(Intercept)]", "(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice")
)
```

## Prediction Output

Our analysis aims to predict the price trends of Loblaws' "No Name" products from November 20, 2024, to February 2025. The results of our model are illustrated in @fig-results1 and @fig-results2. @fig-results1 presents the predicted trends categorized by product type and size, providing a comparison across different products. The model predicts overall moderate price stability, with some variation between product types and sizes. @fig-results2 shows the prediction distribution categorized by product size.

At the product category level, the left panel of @fig-results1 shows that the price of Snacks remains relatively stable, while the other four categories exhibit comparatively larger fluctuations. These trends align with historical data. However, compared to historical data, the model underestimates the price volatility in certain categories. For instance, the historical data shows significant short-term price fluctuations in Seasoning, which are not fully reflected in the predicted results.

For product size, the right panel of @fig-results1 and @fig-results2 both indicate clear price stratification among Small, Large, and Extra Large products, with larger-sized products having higher average prices. The model successfully replicates a reasonable stratification structure, but the predicted price range is narrower, and the price volatility is lower compared to the historical observations.

```{r, fig.width=8, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-results1
#| fig-cap: Prediction of average price trend for Loblaws' No Name Grocery by type and by size
#| warning: false

# Summarize data by `type` and `date`
type_trends <- prediction_data %>%
  group_by(date, type) %>%
  summarize(average_price = mean(adjusted_price, na.rm = TRUE))

p5 <- ggplot(type_trends, aes(x = date, y = average_price, color = type)) +
  geom_line(size = 1) +
  labs(
    title = "Predicted Price Trends Over Time by Product Type",
    x = "Date",
    y = "Average Predicted Price (CAD)",
    color = "Product Type"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

# Summarize data by `size` and `date`
size_trends <- prediction_data %>%
  group_by(date, size) %>%
  summarize(average_price = mean(adjusted_price, na.rm = TRUE))

p6 <- ggplot(size_trends, aes(x = date, y = average_price, color = size)) +
  geom_line(size = 1) +
  labs(
    title = "Predicted Price Trends Over Time by Product Size",
    x = "Date",
    y = "Average Predicted Price ($)",
    color = "Product Size"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

p5 + p6
```

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-results2
#| fig-cap: Prediction of average price distribution for Loblaws' No Name Grocery by size
#| warning: false

# with `product_name`, `adjusted_price`, and credible intervals (lower_ci and upper_ci)
ggplot(prediction_data, aes(x = size, y = adjusted_price)) +
  geom_boxplot(aes(lower = adjusted_lower_ci, upper = adjusted_upper_ci), color = "blue", fill = "tomato", alpha = 0.5) +
  geom_hline(yintercept = mean(prediction_data$adjusted_price), linetype = "dashed", color = "black") +
  labs(
    title = "Prediction of Average Product Prices by Product Name",
    x = "Product Name",
    y = "Predicted Price ($)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

```

# Discussion{#sec-discussion}

## Price Forecast

In this paper, we focus on forecasting the price trends of Loblaws "No Name" products from November 20, 2024, to February 15, 2025. The analysis emphasizes the dynamics of prices categorized by product type and size, identifying behavioral patterns and differences between categories. We incorporated historical data like in @fig-1 and @fig-2 to ensure that the predictions align with observed past trends. Additionally, we designed the prediction dataset with constraints to simulate reasonable price changes—prices adjust at least every 14 days, with fluctuations ranging between 10% and 30% of the previous price. The results in @sec-result, as illustrated in @fig-results1, indicate that while the prices of Loblaws "No Name" products are expected to vary over time, the overall increase or decrease remains relatively limited. Both the predicted and historical data demonstrate similar trends across different categories. This suggests that retail pricing adjustments and predictions for similar product categories may follow patterns that are relatively consistent over time.

From the projected price fluctuations and distributions, the forecast suggests that there will be no significant price increases for Loblaws "No Name" products analyzed in this study from late November 2024 to mid-February 2025. This aligns with Loblaws' stated commitment to "maintain price stability" for No Name products whenever possible.

## Limitations of Existing Data

This section discusses the limitations of the data, data processing, and the proposed model, aiming to assess its robustness and the challenges in capturing the complexities of price changes. Each step, from data collection to model construction, involves certain assumptions and potential biases. Below is a detailed discussion of these aspects:

Firstly, the raw data used for observation and modeling is extensive and complex, with numerous duplicates and classification errors, as detailed in [Appendix -@sec-data-details]. One of the major challenges for the model is providing accurate predictions under limited accuracy and highly complex categorization. The complexity of the data makes it challenging for the model to provide clear interpretations. For example, the model might struggle to distinguish whether price fluctuations are caused by random events (such as promotions) or long-term trends (such as seasonal demand changes). This reduces the model's practical applicability, especially when decision-makers need straightforward conclusions.

The high variability in the data means that the model may struggle to accurately capture key trends and patterns. With a high level of noise, parameter estimates may be significantly affected, leading to wider prediction intervals and reduced reliability. For example, frequent fluctuations or outliers in the price data for certain products could cause the model's mean predictions to deviate from actual median trends. Misclassified product prices make it difficult for the model to correctly assess the impact of product characteristics on price. This may lower the model's ability to differentiate between categories, resulting in predictions that deviate from reality.

To reduce bias, we opted to clean duplicates and classification errors before modeling, attempting to standardize the data format. However, this greatly reduced the usable data volume. Unfortunately, since the true state of the data is unknown, the cleaned data may still contain inaccuracies. In the future, when faced with highly variable and complex categories, the model may need to incorporate additional random effects or interaction terms to account for data heterogeneity. However, such complexity may lead to overfitting, where the model fits the training data too closely but performs poorly on new data. This might also increase computational demands to impractical levels.

@loblaws2022response also emphasized the importance of regional and independent retailers in maintaining market competitiveness, particularly in non-urbanized areas. However, due to the dataset from @hammer only include the data from one community in Canada, our model cannot track the price behavior across regions, which may oversimplify the actual dynamics of the Canadian grocery market. If geographic information were incorporated, introducing regional effects or hierarchical modeling could enhance prediction accuracy, especially in capturing price trend differences between urban centers and less competitive markets.

## Real-World Impact

Another limitation of our analysis lies in its reliance on historical pricing data as the primary input. While this approach ensures consistency, it risks overlooking real-time market signals, such as sudden shifts in consumer preferences due to trends, unexpected supply chain disruptions for certain goods, or the introduction of new products with prices significantly different from similar items in the same category. This can result in predictions lagging behind actual market conditions, leading to inaccuracies in forecasting price volatility and trends and diminishing the model's real-time applicability. Additionally, the model does not account for external factors like inflation, competitor pricing, or macroeconomic indicators. These variables also influence retail prices, but due to a lack of data, they were excluded. Furthermore, the relatively narrow predicted price range suggests that the model may rely too heavily on prior assumptions, reducing its sensitivity to changes in price dynamics.

In future research, if sufficient data becomes available, we plan to enhance the model by incorporating dynamic factors or real-time updating mechanisms to respond quickly to market changes. This could include forecasting pricing for new products and simulating their potential impact on the market. However, like mentioned in @loblaws2022response price categorization in retail is often determined by consumer demand, packaging specifications, or branding strategies, which may not directly align with physical dimensions or averages of the products. This challenge is difficult to resolve, as we cannot predict changes in consumer preferences, meaning the results may not fully reflect actual market characteristics. The predictions will always lag behind real-world outcomes, with the goal being to move as close to reality as possible.

## Loblaws' Pricing Strategy

In @walrus2022grocer, the article highlights fairer supplier agreements, with one of the core goals of the code of conduct being to reduce conflicts and non-transparent practices between suppliers and retailers. This could provide greater market access for small and medium-sized suppliers, increasing the flexibility of private label supply chains, and would enable Loblaws to maintain price stability at lower supplier costs [see more details in @walrus2022grocer]. Combined with the low price volatility observed in our prediction results, this dynamic could further support Loblaws in offering competitive prices through its "No Name" brand, as these products primarily emphasize affordability. The code of conduct could help Loblaws maintain or improve No Name product quality without significantly increasing prices by reduced costs.

Based on the content from @canadiangrocer2023loyalty and @Li_Sexton_Xia_2006, we suggest that Loblaws positions its private label "No Name" at a lower price point to attract price-conscious consumers while gaining stronger bargaining power within the supply chain. This strategy likely allows Loblaws to stabilize or even reduce supplier costs while maintaining customer satisfaction with pricing. Furthermore, promotional activities such as discounts and loyalty rewards significantly influence consumer purchasing behavior and brand preferences. This aligns with Loblaws' approach to engaging consumers through its PC Optimum program.

Moreover, by using the extensive consumer data collected through PC Optimum, Loblaws can more accurately adjust product pricing and promotional strategies, continually refining its market positioning. Loyalty programs not only attract price-sensitive consumers but also increase the switching costs for customers moving to competitors, further boosting market share [see more details in @canadiangrocer2023loyalty]. Thus, the conjecture is, even if prices for other goods rise, Loblaws can mitigate negative consumer reactions by offering discounts, reward points, and stable pricing for its private label products. The "No Name" brand helps enhance consumer tolerance for price increases by maintaining an image of affordability. By preserving the perception of low prices, Loblaws reduces the visibility of price hikes, allowing it to sustain its market position and profitability without losing its overall customer base.

However, based on the report by @mcdonald-gibson_outraged_2024, Loblaws’ policies have not effectively appeased consumer sentiment. Consumers believe that Loblaws’ record profits amid rising food prices demonstrate corporate greed and disregard for consumer interests. Although the boycott has gained some attention, Loblaws’ stock price continues to rise, showing that the boycott has had limited financial impact on the company. When it comes to groceries, consumers have relatively limited choices. They usually choose stores that are convenient for their daily shopping needs. Furthermore, as a large corporation with multiple subsidiaries, such as Shoppers, it is challenging for consumers to completely boycott or significantly reduce their shopping frequency without affecting their daily lives. We believe that companies like Loblaws should soon make a balance between service, procurement costs, and pricing, minimizing unnecessary expenses and providing consumers with affordable groceries.

\newpage

\appendix

# Appendix {-}

# Data Collection Methodology Overview and Evaluation {#sec-data-details}

Observational data forms the core of our models; however, unlike experimental data, it is inherently limited by sampling biases, incomplete data, misclassification or duplicate classification of products, and inconsistent cross-product size standardization. Imagine you are assembling a complex machine, where each component must fit perfectly to deliver precision and efficiency. This is precisely how the Hammer Project operates—piecing together grocery pricing data to construct a unified picture of competition and pricing behaviors in Canada’s grocery market. In this section, we explore the foundation of Hammer Project’s data collection and analysis methodologies.

## Data Sources and Collection
The [Hammer Project](https://jacobfilipp.com/hammer/) collects its data through web scraping, capturing pricing information directly from the user interfaces of major grocery retailers’ websites without using APIs. Prices are specifically for the "in-store pickup" option and are based on a neighborhood in Toronto. The dataset includes eight vendors: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods.

However, relying on web scraping instead of APIs comes with limitations. APIs often provide more accurate and reliable historical price data, while web scraping can only capture data currently displayed on the website. For example, if a product is removed from the webpage after selling out rather than being marked as "out of stock," it may not be captured. Changes in website structure or server interruptions can also lead to inconsistencies, requiring researchers to perform additional data cleaning and supplementation. Furthermore, focusing solely on "in-store pickup" prices, without considering delivery or other purchasing methods, might fail to accurately reflect overall pricing strategies or market behaviors of grocers. This limitation narrows the dataset’s applicability. Additionally, collecting data from a single geographic area may restrict the generalizability of findings to other regions of Canada.

To ensure comprehensive coverage of grocery price trends, the dataset expanded its scope after the initial phase. Early data collection (from February 28 to July 10/11) focused on a small subset of products, but a broader variety of products was included starting July 10/11. This later expansion better reflects the diversity of products and pricing behaviors across vendors.
However, this sudden introduction of new data after July 10/11 poses challenges for analyzing price trends over time. It could create disruptions within categories, impacting the consistency of earlier patterns. Since the February to July period constitutes a significant portion of the current dataset (as of November 2024), researchers cannot exclude this period entirely to eliminate disturbances.

## Data Cleaning and Integration
@hammer processes its data into two core tables: a product table containing metadata and a raw table with time-series price data. These tables are joined using unique product identifiers. Challenges include managing duplicate entries for the same product and distinguishing between similar items. As noted on the Hammer Project website, a single product might appear multiple times under different categories, or subtle branding differences might complicate matching. Addressing these issues requires future data cleaning. However, since the data cannot be revisited retrospectively to correct inaccuracies from the time of collection, these challenges make ensuring complete data reliability more difficult.

Additionally, the Hammer Project dataset encourages users to explore targeted research questions. Examples include:

- `How do prices for common grocery bundles (e.g., eggs, milk, and bread) change over time and across vendors?`

- `Are "sales" genuine discounts, or do they follow price hikes?`

- `Which vendors consistently offer the lowest prices across comparable products?`

These questions align with the structured approach seen in polling methodology, where clear and purposeful design guides analysis. The transparency and flexibility of the dataset support academic and legal research, fostering a deeper understanding of Canada’s grocery industry.

# Model details {#sec-model-details}
R package `bayesplot` [@bayesplot] is used for posteriors.

Due to space limitations, we have chosen to display only the posterior and prior comparisons for a subset of variables in the posterior analysis. If we were to display all variables included in the model, the chart would become overly complex and difficult to interpret. Although the effects of multiple variables differ, focusing on this subset may give the impression that the model’s sensitivity to variations across different factors is limited. Thus, we have selected representative variables, including the intercept, size categories (`sizeLarge` and `sizeSmall`), and product types (`typeSnacks`, `typeFlour/Rice`, and `typeSeasoning`). The output results for other variables not shown are generally consistent with the trends observed here. Additionally, in regression models (such as stan_glmer), categorical variables are typically automatically converted into dummy variables. One category is set as the reference category (baseline), and its effect is included in the intercept, while the coefficients for other categories represent their offsets relative to the baseline. For example, in our model, if `Extra Large` is set as the baseline (default), it will not appear as a separate coefficient. The coefficients for `sizeLarge` and `sizeSmall` would then represent their deviations relative to `Extra Large`.

## Posterior predictive check
@fig-post1 displays a comparison between the distribution of observed data $y$ and the distribution of simulated data $y_{rep}$ generated by our model. The dark solid line represents the observed data's actual distribution, while the multiple light lines correspond to distributions from various simulated datasets. If the distribution of simulated data $y_{rep}$ closely resembles that of the observed data $y$, it suggests that the model effectively captures the underlying characteristics of the observed data.

In @fig-post1, we observe that the simulated data distributions (light lines) align closely with the real data distribution (dark line) across the entire range, particularly at the peaks and troughs of the observed data. This indicates that the model has captured the distributional features of the observed data, including its skewness and multimodal structure. The ability to replicate these features demonstrates the model's strong predictive capacity and its reliability in estimating posterior distributions. This result builds confidence in the model's suitability for further inference and prediction.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post1
#| fig-cap: PP Check for Model
#| warning: false

pp_check(first_model)
```

@fig-post2 displays the MCMC sampling trace plots for six key parameters in our model, including the `(Intercept)`, size variables `sizeLarge` and ` sizeSmall`, and type variables `typeSnacks`, `typeFlour/Rice` and `typeSeasoning`. Each plot includes four chains from Chain 1 to Chain 4, representing independent MCMC sampling processes. In a well-converged model, our expectation is that each chain to stabilize and fluctuate around a central value without showing a clear trend (i.e., the mean of the samples converges over time). Additionally, the chains for a parameter should exhibit consistent behavior, indicating agreement among the chains.

From @fig-post2, we observe that the sampling chains for all six parameters show a steady state, with fluctuations around a central value and no observable trends over the iterations. Moreover, the chains for each parameter are closely aligned, demonstrating consistency between the chains. This isuggests that the model has reached convergence and the posterior distributions for these parameters are stable. These trace plots support the reliability of our MCMC sampling process and provide confidence in the validity of the posterior estimates.

```{r, fig.width=7, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post2
#| fig-cap: MCMC sampling trace plots for model
#| warning: false

plot(first_model, "trace", pars = c("(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice", "typeSeasoning"))
```
@fig-post3 shows the $\hat{R}$ diagnostic statistic for our model, which is used to assess the convergence of MCMC sampling. Ideally, is that the $\hat{R}$ value would be close to value 1, which indicates consistent variance across different MCMC chains, signifying that the model has reached a stable state. The horizontal axis represents the $\hat{R}$ values, where values closer to 1 indicate better convergence. Our expectation is that, all parameters would have $\hat{R}$ values below 1.05, which means minimal variance differences between chains and consistency across them. This plot shows that all $\hat{R}$ values are concentrated around 1, and all $\hat{R}$ values are lower than the threshold of 1.05 (indicated by the vertical dashed line on the right), suggesting that the sampling chains have converged well, providing thw reliability for our model results.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post3
#| fig-cap: R hat Diagnostic statistic for model
#| warning: false

plot(first_model, "rhat")
```

@fig-post4 shows the posterior distributions of certain parameters in our model, including `(Intercept)`, size variables `sizeLarge` and `sizeSmall`, and type variables `typeSnacks`, `typeFlour/Rice`, and `typeSeasoning`. The shape and location of these distributions help us understand the extent to which the observed data supports these parameter estimates and the degree of uncertainty associated with them.

The `(Intercept)` is centered nearly around 2.5, reflecting the baseline level for product prices in our model when all other factors are held constant. The narrow and peaked distribution suggests strong support for this estimate with relatively low uncertainty. For `sizeLarge` and `sizeSmall`, the distributions are shifted slightly to the left, indicating that these sizes are generally associated with lower prices compared to the baseline category `ExtraLarge`. Type variables `typeSnacks`, `typeFlour/Rice`, and `typeSeasoning` exhibit posterior distributions centered close to zero, implying that their influence on product pricing is minimal or uncertain, as the data does not strongly support a directional impact.

Overall, the results suggest that the model successfully captures the baseline pricing level and the relative influence of product sizes, but it struggles to discern clear effects for product types, possibly due to overlapping pricing patterns or insufficient variation within these categories.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post4
#| fig-cap: posterior distributions of four parameters in model
#| warning: false

plot(first_model, "areas", pars = c("(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice", "typeSeasoning"))
```

\newpage

# References